# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r echo=TRUE}
rawdata <- read.csv("activity.csv")
cleandata <- subset(rawdata, complete.cases(rawdata))
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
### use a data.frame with the number of rows equal to number of days
### name the rows according to days
### name a collumn as "sum_of_seps"
m <- data.frame(sum_of_seps = 1:length( levels( cleandata$date ) ), 
                row.names = levels( cleandata$date ) )

### save in m sum of steps per day
for (i in levels(cleandata$date)) m[i,1] <- sum( cleandata$steps[ cleandata$date == i] )

### "Make a histogram of the total number of steps taken each day" *quote from assignment
### take the default binwidth that is range/30 (to skip warning message)
library(ggplot2)
qplot(sum_of_seps, data = m, binwidth = ( max( m$sum_of_seps ) - min( m$sum_of_seps ) )/30 )

### "Calculate and report the mean and median total number of steps taken per day" *quote from assignment
smean <- mean( m[,1] )
smedian <- median( m[,1] )
```
The **mean** of total number of steps taken per day is `r smean`.  
and **median** of total number of steps taken per day is `r smedian`.

## What is the average daily activity pattern?
```{r echo=TRUE}
### t is a vector of intervals of 5 minutes
t <- seq (from = min(cleandata$interval), to = max(cleandata$interval), by = 5)

### use a data.frame with the number of rows equal to number of intervals
### "interval_no" column contains numner of intervals
### "avg_of_seps" column is for number of steps per interval averaged across all days
p <- data.frame(interval_no = t, avg_of_seps = 1:length(t))

### save in p average of steps per interval across all days
for (j in 1:length(t) ) p[j,2] <- mean( cleandata$steps[ cleandata$interval == t[j] ] )

### not every interval was recordered
p_clean <- subset(p, complete.cases(p))

### "Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)" *quote from assignment
        
```


## Imputing missing values
```{r echo=TRUE}

```


## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}

```
