# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r echo=TRUE}
rawdata <- read.csv("activity.csv")
cleandata <- subset(rawdata, complete.cases(rawdata))
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
### use a data.frame with the number of rows equal to number of days
### name the rows according to days
### name a collumn as "sum_of_seps"
m <- data.frame(sum_of_seps = 1:length( levels( cleandata$date ) ), 
                row.names = levels( cleandata$date ) )

### save in m sum of steps per day
for (i in levels(cleandata$date)) m[i,1] <- sum( cleandata$steps[ cleandata$date == i] )

# sum( subset(cleandata, date == i, select = steps)

### "Make a histogram of the total number of steps taken each day" *quote from assignment
### take the default binwidth that is range/30 (to skip warning message)
library(ggplot2)
qplot(sum_of_seps, data = m, binwidth = ( max( m$sum_of_seps ) - min( m$sum_of_seps ) )/30 )

### "Calculate and report the mean and median total number of steps taken per day" *quote from assignment
smean <- mean( m[,1] )
smedian <- median( m[,1] )
```
The **mean** of total number of steps taken per day is **`r smean`**.  
and **median** of total number of steps taken per day is **`r smedian`**.

## What is the average daily activity pattern?
```{r echo=TRUE}
### t is a vector of intervals of 5 minutes
t <- seq (from = min(cleandata$interval), to = max(cleandata$interval), by = 5)

### use a data.frame with the number of rows equal to number of intervals
### "interval_no" column contains numner of intervals
### "avg_of_seps" column is for number of steps per interval averaged across all days
p <- data.frame(interval_no = t, avg_no_of_seps = 1:length(t))

### save in p average of steps per interval across all days
for (j in 1:length(t) ) p[j,2] <- mean( cleandata$steps[ cleandata$interval == t[j] ] )

### not every interval was recordered
p_clean <- subset(p, complete.cases(p))

### "Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)" *quote from assignment
library(lattice)
xyplot( avg_no_of_seps ~ interval_no, data = p_clean, type = "l" )

### "Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?"
max_int <- p_clean[,1][ p_clean[,2] == max( p_clean[,2] ) ] 
```
The 5-minute interval number **`r max_int`** contains the maximum number of steps on average across all the days in the dataset

## Imputing missing values
```{r echo=TRUE}
### "Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)" *quote from assignment
na_no <- nrow(rawdata) - nrow(cleandata)
```
The total number of missing values in the dataset (i.e. the total number of rows with NAs) is **`r na_no`**

```{r echo=TRUE}
### "Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc." *quote from assignment
### "Create a new dataset that is equal to the original dataset but with the missing data filled in." *quote from assignment
restoreddata <- rawdata
for (k in 1:nrow(restoreddata)) {
        if ( is.na( restoreddata[k,"steps"] ) ) {
                restoreddata[k,"steps"] <- p_clean[,2][ p_clean[,1] == restoreddata[k,"interval"] ]
        }
}
```

### A histogram of the dataset with the missing data filled in by the average values per interval on average across all the days in the dataset 

```{r}
### "Make a histogram of the total number of steps taken each day" *quote from assignment

### use a data.frame with the number of rows equal to number of days
### name the rows according to days
### name a collumn as "sum_of_seps"
m_new <- data.frame(sum_of_seps = 1:length( levels( restoreddata$date ) ), 
                row.names = levels( restoreddata$date ) )

### save in m_new sum of steps per day
for (l in levels(restoreddata$date)) m_new[l,1] <- sum( restoreddata$steps[ restoreddata$date == l] )

### "Make a histogram of the total number of steps taken each day" *quote from assignment
### take the default binwidth that is range/30 (to skip warning message)
library(ggplot2)
qplot(sum_of_seps, data = m_new, binwidth = ( max( m_new$sum_of_seps ) - min( m_new$sum_of_seps ) )/30 )

qplot(sum_of_seps, data = m, binwidth = ( max( m$sum_of_seps ) - min( m$sum_of_seps ) )/30 )

### "Calculate and report the mean and median total number of steps taken per day" *quote from assignment
smean_new <- mean( m_new[,1] )
smedian_new <- median( m_new[,1] )
dif_mean <- smean - smean_new
dif_median <- smedian - smedian_new
```
The **mean** of total number of steps taken per day is **`r smean_new`**.  
and **median** of total number of steps taken per day is **`r smedian_new`**.

```{r}
### "Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?" *quote from assignment" *quote from assignment
```
The **mean** of total number of steps taken per day (from the estimates from the first part of the assignment) was **`r smean`**, the same figure with the restored NAs by average numbers is **`r smean_new`**.  
The **median** of total number of steps taken per day (from the estimates from the first part of the assignment) was **`r smedian`**, the same figure with the restored NAs by average numbers is **`r smedian_new`**.

The impact (difference) in **mean** value is **`r dif_mean`** and in **median** is **`r dif_median`**. The impact is not negligible.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}

```
